---
title: "Customising functionalities"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customising functionalities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(mcmcglm)
withr::local_seed(42)
```

## Adding methods for new families

the only missing piece for any other families is a method for calculating the log-density of the distribution, which has been made very easily 

Fx. for inverse gaussian distribution add the method
```{r, eval = FALSE}
#' @export
log_density.inverse.gaussian <- function(family, main_parameter, Y, ...) {
  statmod::dinvgauss(Y, mean = main_parameter, ..., log = T)
}
```

and when calling `mcmcglm` remember to add the additional arguments to this density through the `log_likelihood_extra_args` argument in this case by doing fx.
```{r, eval = FALSE}
mcmcglm <- mcmcglm(family = "inverse.gaussian",
                   log_density_extra_args = list(shape = x, dispersion = y),
                   "other arguments")
```

## Using other samplers

