---
title: "Adding methods for families"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding methods for families}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
```

```{r setup, include = FALSE}
library(mcmcglm)
withr::local_seed(42)
```

## Adding methods for new families

It is very, very easy to extend the package to include a family that is not existing in the package. The only missing piece for any other families is a method for calculating the log-density of the distribution.

Since the calculation of the log-density for a family is implemented using S3 methods, all the user needs to do is define a function called `log_density.*family_name*` with *family_name* specified as the relevant family.

Below is an example showcasing the single line of code a user needs to write to enable the use of an inverse gaussian family

### Example of implementing method for inverse gaussian family

First, we generate some data to showcase the example.
```{r, label = data-generation}
n <- 1000
x1 <- rexp(n, 2)
x2 <- rbinom(n, 1, .5)
b0 <- 1
b1 <- 1.5
b2 <- 2
lin_pred <- b0+b1*x1+b2*x2

invgauss_fam <- inverse.gaussian()
y_invnorm <- statmod::rinvgauss(n, mean = invgauss_fam$linkinv(lin_pred),
                                shape = 1, dispersion = 1)
dat_invnorm <- data.frame(Y = y_invnorm, X1 = x1, X2 = x2)
```

All the user needs to do is specify a method for computing the density of an inverse gaussian distribution like so:
```{r define-invgauss}
log_density.inverse.gaussian <- function(family, mu, Y, ...) {
  statmod::dinvgauss(Y, mean = mu, ..., log = T)
}
```

Then, `mcmcglm()` can be called like so:
```{r, eval = FALSE}
invnorm <- mcmcglm::mcmcglm(formula = Y ~ .,
                            family = inverse.gaussian(),
                            log_likelihood_extra_args = list(shape = 1, dispersion = 1),
                            data = dat_invnorm,
                            beta_prior = distributional::dist_gamma(1, 1),
                            w = 0.5)
trace_plot(invnorm)
```

where we specify `inverse.gaussian()` as the `family` argument, and we make sure to add the additional arguments to this density through the `log_likelihood_extra_args` argument.
